Group 5 Report

### 2022-08-02 Report/Group 5 report.Rmd
Adding dashboard report code. Editing throughout.

---
title: "Group 5 Report"
author: "ISSSV1337 Group 5"
date: "2022-08-03"
output:
  pdf_document: 
    toc: true
    highlight: pygments
    extra_dependencies: "subfig"
  html_document: default
---

@@

We discussed creating a webcrawler that could collect data across
different websites, but quickly decided that this would be too
difficult, and that it would only work with selected organizations. The next
suggested solution was to create code that could read PDF documents
that OPX downloaded themselves. Our code would analyse and organise the
uploaded data from the PDF documents.

@@

- Financial statements in annual report PDF documents. 
- Webscrape Proff.no for financial statements. 
- Webscrape the Brønnøysund registry.

@@

The OPX case requested two tasks -- webscrape financial data, and create
KPIs. During phase three, we scrapped all ideas and progress from the
first two stages, and focused on creating KPIs.

@@

We wanted to use webscraping to collect the data, so that we could have
automated the data input process. As described, we explored different
potential solutions. We experimented on coding such that data could have
been imported from XML files. However, once we realised that the
financial data was of little use, this experimentation stopped.

@@

We copied these financial statements manually into an Excel document.
Perhaps a "simple" solution, however our code requires OPX to import
both an Excel document themselves into shiny, as well as a HTML document
from Proff.no.

@@

@@

As the dashboard uses a fair number of packages, it uses the pacman package to ensure they are installed, to avoid the user being confronted with messages about missing packages:

@@

The sidebar includes file upload boxes for Excel and html files, numeric input boxes in which the user can input the years covered by the Proff.no data, and an 'Analyse' button which is connected to two Shiny `observe() |> bindEvent()` structures, ensuring that most of the code is not executed until the user has finished entering data. (Though both structures await the same button click, the division was necessary to remain within R's 10,000 byte limit.)

@@

`r if (knitr::is_latex_output()) line`
```{r dashboard_no_information, eval=FALSE}
## no_information plot ---------------------------------------------------------
# Function to print a 'no information' plot, if necessary
no_information <- function() {
  empty_plot("No information available")
}

## empty_plot-------------------------------------------------------------------
# Function to print an empty plot with a supplied message
empty_plot <- function(warning) {
  ggplot() +
    # text geom with a space to create the plot, but otherwise show nothing
    geom_text(aes(0, 0, label = " ")) +
    # as blank a slate as possible
    theme_void() +
    annotate(
      "text",
      label = warning,
      # Using the same palette as the other non-profit/loss plots
      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
    )
}
```

@@

`r if (knitr::is_latex_output()) line`
```{r}
zero_data <- function() {
  empty_plot(
    "All values are zero or missing.
    This is unusual and may
    indicate a data problem.")
}
```

@@

Both plots feature a certain amount of data transformation, including renaming the variables, so that the legend includes appropriate information; pivoting - in order to use the `fill` aesthetic; and ordering of the categories, so that they occur in an appropriate order. Excel data are presumed to be supplied in NOK.

@@

All the plots based on the Excel data make use of the colour-blind-friendly "Hokusai2" palette from the MetBrewer package; in the case of the revenues plot this is not done using the convenience `scale_fill_met_d()` function but rather the `met.brewer()` function in conjunction with `scale_fill_manual()` as the base palette only contains six colours, but the plot has seven categories. The same function is used to generate the text colour in the 'no information' and 'bad data' plots.

@@

@@

@@

The limitation with the accessible data is first and foremost the lack 
of relevant data in the annual reports from the NGOs.  Since there is 
no obligation to provide annual accounts, they often have more information 
about the organization in general rather than numbers. If they do provide 
the annual accounts, they are not standardized. This means that different 
NGOs do not have an established template for what the different posts in 
the annual accounts should contain.

Also, these annual reports themselves appear in different formats. If they 
do appear in the same format, the accounts we want to retrieve are often 
in an image format rather than text. The lack of data, standardization and 
diverse formats made it impossible to create an automated solution for 
collecting data based on the NGOs’ annual reports.

The limitation with the non-accessible data is that the data OPX needs to 
calculate the different KPIs do not exist in the annual reports.  We 
experimented with an annual report from Brønnøysundregisteret. Although 
the reports were standardized, the different posts in the report did not 
match the template from OPX. Additionally the data that OPX is interested 
in and which is necessary for calculating the relevant KPIs are not 
available.

Also, with limited experience with KPIs it was hard to sort out which ones 
that would be of relevance and not. After examining the different 
alternatives, we concluded that the NGOs could report in a standardized 
excel document from which our code retrieves the data to calculate the KPIs.

In addition, we looked at the draft of a new law on associations from 
2021 by the Ministry of Trade, Industries, and Fisheries, to see if it 
contained changes in the laws regarding annual reports, which would have 
made the data needed available in the future. This turned out to be a dead 
end, as the draft focused largely on smaller and local associations and 
contained no changes in reporting obligations. Therefore, if this law were 
to be implemented, it would not be of any relevance to OPX. As a result, 
due to OPX’s need for this data to be available, we recommend discussions 
with relevant politicians and ministry officials as a starting point.

Based on the findings of working with this case, we would recommend looking 
at best practices for further development. One good example could be charity 
navigator.

@@

In this last section, we would like to describe how we experienced this process and what we have learned. As explained, the main challenge with the task was highlighted by an unsolvable due to non-existent data. In cases where data were available, they were inadequate.

@@

The lectures allowed us to learn about many interesting topics. However, as the connection between what we learned there and what we got to utilize in the case was distant, and parts were too advanced for some of us, especially those without any particular experience within statistics.

Despite this, we have done our best to work with the resources at hand. As shown in this presentation, trying out solutions and later scrapping them has been a big part of the problem-solving process. We also got an insight into real-life work situations, such as the challenges that come when what is desired clashes with what is possible. We are certain that these experiences will be useful for all of us in the future, and we are thankful for being given this opportunity.

>>>>

---
title: "Group 5 Report"
author: "ISSSV1337 Group 5"
date: "2022-08-03"
output:
  pdf_document: 
    toc: true
    highlight: pygments
    extra_dependencies: "subfig"
---

@@

<!--# En webcrawler er en bot -->

We discussed creating web scraping code that could collect data across
different websites, but quickly decided that this would be too
difficult, and that it would only work with selected organisations. The next
suggested solution was to create code that could read PDF documents
that OPX downloaded themselves. Our code would analyse and organise the
uploaded data from the PDF documents.

@@

- Financial statements in annual report PDF documents. 
- Web scrape Proff.no for financial statements. 
- Web scrape the Brønnøysund registry.

@@

The OPX case requested two tasks -- web scrape financial data, and create
KPIs. During phase three, we scrapped all ideas and progress from the
first two stages, and focused on creating KPIs.

@@

We wanted to use web scraping to collect the data, so that we could have
automated the data input process. As described, we explored different
potential solutions. We experimented on coding such that data could have
been imported from XML files. However, once we realised that the available
financial data was of little use, this experimentation stopped.

@@

We copied these financial statements manually into an Excel document.
Perhaps a 'simple' solution, however our code requires OPX to import
both an Excel document themselves into shiny, as well as a HTML document
from Proff.no.

@@

<!--# Fra Visualisations (nedenfor): The plots based on data from Proff.no often demonstrate surprising lacunae, but it difficult to know where they originate -- NGO reporting, Brønnøysund, or Proff.no.-->

@@

As the dashboard uses a fair number of packages, it uses the pacman package to ensure they are installed, to avoid the user being confronted with (cryptic) messages about missing functions:

@@

The sidebar includes file upload boxes for Excel and html files, numeric input boxes in which the user can input the years covered by the Proff.no data (needed because of the highly variable availability of historical data), and an 'Analyse' button which is connected to two Shiny `observe() |> bindEvent()` structures, ensuring that most of the code is not executed until the user has finished entering data. (Though both structures await the same button click, the division was necessary to remain within R's 10,000 byte limit.)

@@

`r if (knitr::is_latex_output()) line`
```{r dashboard_no_information, eval=FALSE}
# Function to print a 'no information' plot, if necessary
no_information <- function() {
  empty_plot("No information available")
}

# Function to print an empty plot with a supplied message
empty_plot <- function(warning) {
  ggplot() +
    # text geom with a space to create the plot, but otherwise show nothing
    geom_text(aes(0, 0, label = " ")) +
    # as blank a slate as possible
    theme_void() +
    annotate(
      "text",
      label = warning,
      # Using the same palette as the other non-profit/loss plots
      x = 0, y = 0, size = 8, colour = met.brewer("Hokusai2", 1)
    )
}
```

@@

`r if (knitr::is_latex_output()) line`
```{r}
zero_data <- function() {
  empty_plot(
    "All values are zero or missing.
    This is unusual and may
    indicate a data problem."
  )
}
```

@@

Both plots feature a certain amount of data transformation, including renaming the variables, so that the legend includes appropriate information; pivoting -- in order to use the `fill` aesthetic; and ordering of the categories, so that they occur in an appropriate order. Excel data are presumed to be supplied in NOK.

@@

All the plots based on the Excel data make use of the colour-blind-friendly 'Hokusai2' palette from the MetBrewer package; in the case of the revenues plot this is not done using the convenience `scale_fill_met_d()` function but rather the `met.brewer()` function in conjunction with `scale_fill_manual()` as the base palette only contains six colours, but the plot has seven categories. The same function is used to generate the text colour in the 'no information' and 'bad data' plots in order to present the user with a somewhat unified plot aesthetic.

@@

`r if (knitr::is_latex_output()) line`
```{r dashboard_download, eval=FALSE}
output$report <- downloadHandler(

  # creating the file name using the NGO title and system date
  filename = function() {
    str_c(title(), "_", Sys.Date(), ".pdf")
  },
  content = function(file) {
    # Based on code in Mastering Shiny -- creating a temporary directory and
    # file
    tempReport <- file.path(tempdir(), "report.Rmd")
    file.copy("report.Rmd", tempReport, overwrite = TRUE)

    # parameters passed to the R Markdown document
    params <- list(
      htmlfile = input$html$datapath,
      excelfile = input$excel$datapath,
      start = input$start,
      end = input$end
    )

    # notification -- disappears when report generation is complete
    id <- showNotification(
      "Generating report...",
      duration = NULL,
      closeButton = FALSE
    )
    on.exit(removeNotification(id), add = TRUE)

    # R Markdown render
    rmarkdown::render(tempReport,
      output_file = file,
      params = params,
      envir = new.env(parent = globalenv())
    )
  }
)
```

@@

`r if (knitr::is_latex_output()) line`
```{r dashboard_download_button, eval=FALSE}
## Download button -----------------------------------------------------------
output$downloadUI <- renderUI({
  if (available_excel() && available_html()) {
    downloadButton("report", "Download report")
  }
})

## Data exists condition testing----------------------------------------------
# Essentially pretty random tests to connect to the individual
# validate() functions upstream in the reactive chain

# Excel
available_excel <- reactive({
  TRUE
}) |>
  bindEvent(setequal(
    names(exceldata()),
    c("year", "firm", "category", "sub_category", "name", "value")
  ))

# HTML
available_html <- reactive({
  TRUE
}) |> bindEvent("lederlonn" %in% names(htmldata()))
```

@@

The limitation with the accessible data is first and foremost the lack 
of relevant data in the annual reports from the NGOs. Since there is often 
no obligation to provide annual accounts, they often have more information 
about the organisation in general rather than numbers. If they do provide 
the annual accounts, they are not standardized. This means that different 
NGOs do not have an established template for what the different posts in 
the annual accounts should contain.

Also, these annual reports themselves appear in different formats. If they 
do appear in the same format, the accounts we want to retrieve are often 
in an image format rather than text. The lack of data, standardisation and 
diverse formats made it impossible to create an automated solution for (safely)
collecting data based on the NGOs’ annual reports.

The limitation with the non-accessible data is that the data OPX needs to 
calculate the different KPIs do not exist in the annual reports.  We 
experimented with an annual report from Brønnøysund. Although 
the reports were standardised, the different posts in the report did not 
match the template from OPX. Additionally the data that OPX is interested 
in and which is necessary for calculating the relevant KPIs are not 
available.

Also, with limited experience with KPIs it was hard to sort out which ones 
that would be of relevance and not. After examining the different 
alternatives, we concluded that the NGOs could report in a standardised 
excel document from which our code retrieves the data to calculate the KPIs.

In addition, we looked at the draft of a new law on associations from 
2021 by the Ministry of Trade, Industries, and Fisheries, to see if it 
contained changes in the laws regarding annual reports, which would have 
made the data needed available in the future. This turned out to be a dead 
end, as the draft focused largely on smaller and local associations and 
contained no changes in reporting obligations. Therefore, if this law were 
to be implemented, it would not be of any relevance to OPX. As a result, 
due to OPX’s need for this data to be available, we weould suggest discussions 
with relevant politicians and ministry officials as a starting point.

Based on the findings of working with this case, we would recommend looking at best practices for further development. One well-known established actor in the field is Charity Navigator.

@@

In this last section, we would like to describe how we experienced this process and what we have learned. As explained, the main challenge with the task was thatb it was in many ways unsolvable due to non-existent data. In cases where data were available, they were inadequate.

@@

The lectures allowed us to learn about many interesting topics. However, as the connection between what we learned there and what we got to utilise in the case was distant, and parts were too advanced for some of us, especially those without any particular experience within statistics.

Despite this, we have done our best to work with the resources at hand. As shown in this presentation, trying out solutions and later scrapping them has been a big part of the problem-solving process. We also got an insight into real-life work situations, such as the challenges that come when what is desired clashes with what is possible. We are certain that these experiences will be useful for all of us in the future, and we are grateful for being given this opportunity.
